---
title: "PEG Africa Case Study Submission by Marvin Adjei Kojo Lomo"
author: "Marvin Adjei Kojo Lomo"
date: "3/25/2021"
output:
  pdf_document: default
  html_document: default
---
# Case Study 1: PEG Africa - Repayment

## Scenario 

PEG Africa's main clients are consumers who lack both access to reliable electricity and any formal banking services (a.k.a the unbanked). The firm offers a sustainable option to meet residential electricity needs via the Solar Home System (three lights, phone and a radio) to consumers (living on $5 - $10). Consumers can build their credit for additional products and services over time.


## Ghana Company Profile

1. 35 Service Centers spread into 7 regions. 
2. Region is managed by an ASM covering about 5 Service Centers.
3. Each Service Center has 4 - 6 DSR managed by the Sales Field Manager.


## Problem Statement

Despite increase in sales over the past three years, reaching 42000 customers across Ghana, there is a high default rate of solar devices. Bad payer behavior has a huge cost for the company. 

# Adidome Service Center in the Volta Region

Portfolio - 198 customers from 2015 - 2017.
Repayment rate - 36%

Task: An understanding of the External Drivers which could explain such a low repayment rate at Adidome compared to the average repayment rate using data.

What could possibly be the situation here? Could it be that generally socio-economic conditions have plummeted so much that consumers can no longer afford PEG Afica's products or the has become an available alternative from competitors or that the National Grid has connected them online now?

To answer this question I used  the Exploratory Data Analysis methodology combined with literature review on loan repayments in Ghana. Due to the paucity of data on District-level demographics, unavailability of Customer-level information at the Adidome Service Center and only 48 Service Center records given, statistical models investigated did not show any significance.

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = '/Users/marvinlomo/Documents/PEG Africa - Case Study')
knitr::opts_chunk$set(echo = TRUE)
```

## Data preparation

```{r}
# Lets load the required libraries
library(tidyverse)
library(readxl)
library(rpart)



# Load the data
repayment_Data <- read_excel("Case 1 Data.xls")


# Converting to tibble Data_frame
repayment_Data <- as_tibble(repayment_Data)

# Lets take a look at the Data
head(repayment_Data)


# Renaming the Repayment Rate column 
repayment_Data$Repayment_rate <- repayment_Data$`Repayment rate`

repayment_Data <- repayment_Data[ , c(1, 2, 4, 5)]


# Converting the region and SC_name to a factor variables

repayment_Data$Region <- as.factor(repayment_Data$Region)
repayment_Data$Name <- as.factor(repayment_Data$Name)

# Focusing in on Adidome_SC
adidome_SC_Data <- repayment_Data %>%
  filter(Region == "Volta South", Name == "Adidome Service Center")
```


## Visualizing the Repayment Data

```{r}
reg_line <- repayment_Data %>%
  select("Sales", "Repayment_rate")




# Investigating Sales vs Repayment rate
repayment_Data %>%
  ggplot(aes(x = Sales , y = Repayment_rate)) + 
  geom_point(aes(size =  Repayment_rate, color = Region, alpha = 2), show.legend = FALSE) +
  geom_smooth(data = reg_line, aes(x = Sales , y = Repayment_rate), 
              method = lm ,se = FALSE, color = "black", lwd = 2)

# Lets look at the correlation between repayment rate with Sales
cor(repayment_Data$Sales, repayment_Data$Repayment_rate)

```
The negative correlation between sales and repayment_rate suggests that Service Centers with high Sales also tend to have a low repayment rate.


## Visualizing the Sales per Region

```{r}

# Investigating average repayment rate at centers vs adidome_SC
repayment_Data %>%
  ggplot(aes(reorder(x = Region, Repayment_rate), y = Repayment_rate, color = Region)) +
  geom_boxplot(show.legend = FALSE) +
  geom_hline(yintercept = mean(repayment_Data$Repayment_rate), lty = 2, lwd = 2, colour = "#333333") +
  geom_point(data = adidome_SC_Data, aes( x = Region, y = Repayment_rate), 
             color = "black", size = 5) +
  labs(x = "Region", y = "Repayment rate") +
  annotate(geom = "text", label = "Average", x = 1.5, y = 0.7, 
           colour = "#1380A1", family = "Helvetica", size = 7) +
  geom_curve(aes(x = 1, y = 0.68, xend = 1, yend = 0.49), colour = "#555555", 
             size = 0.5, curvature = 0.3, arrow = arrow(length = unit(0.03, "npc"))) +
  annotate(geom = "text", label = "Adidome", x = 1.5, y = 0.2, 
           colour = "#1380A1", family = "Helvetica", size = 7) +
  geom_curve(aes(x = 1, y = 0.22, xend = 2, yend = 0.35), colour = "#555555", 
             size = 0.5, curvature = 0.3, arrow = arrow(length = unit(0.03, "npc")))
  


# Investigating average Sales at centers vs adidome_SC
repayment_Data %>%
  ggplot(aes(reorder(x = Region, Sales), y = Sales, color = Region)) +
  geom_boxplot(show.legend = FALSE) +
  geom_hline(yintercept = mean(repayment_Data$Sales), lty = 2, lwd = 2) +
  geom_point(data = adidome_SC_Data, aes(x = Region, y = Sales), 
             color = "black", size = 5) +
  labs(x = "Region", y = "Sales") +
  annotate(geom = "text", label = "Average", x = 1.5, y = 800, 
           colour = "#1380A1", family = "Helvetica", size = 7) +
  geom_curve(aes(x = 1, y = 790, xend = 2, yend = 455), colour = "#555555", 
             size = 0.5, curvature = 0.3, arrow = arrow(length = unit(0.03, "npc"))) +
  annotate(geom = "text", label = "Adidome", x = 5.5, y = 200, 
           colour = "#1380A1", family = "Helvetica", size = 7) +
  geom_curve(aes(x = 5.5, y = 220, xend = 6, yend = 360), colour = "#555555", 
             size = 0.5, curvature = 0.3, arrow = arrow(length = unit(0.03, "npc")))

# It can be seen above that the adidome SC is also selling below the average SC Sales.

```

The gap between the Repayment rate at the Adidome Service Center and the Average 
Service Center repayment rate as seen in the Chart 2 can probably be explained by the 
demographics of the Adidome Area. Various studies on loan repayments conducted have estimated that external drivers that impact loan repayments include 

1. Age
2. Education
3. Gender
4. Household size
5. Loan size
6. Purpose of loan
7. Savings


In the context of the Adidome area, other factors such as whether the individual is connected to the National Grid and Household income can be added to the list of factors that may affect the Repayment rate.

In addition, from a recent report released by the statistical service on Multidimensional Poverty Index of regions, as at 2017, the Volta region along with the three Northern regions were ranked as the poorest regions in the country. This could explain why out of the 10 worst performing Service Centers based on Repayment rate, 5 of the 10 are from the Volta Region.

```{r}
sc_ranked <- repayment_Data %>%
  arrange(Repayment_rate)

head(sc_ranked, 10)
```

Given the above, the following Hypothesis can be constructed: 

1. Null Hypothesis: The above factors do not affect Repayment rate.
2. Alternative Hypothesis: The above factors affect Repayment rate.

Further data collected from customers, and modeling perfomed to understand the factors influencing customer payment behavior.

Should I be given the opportunity to join PEG Africa, I would love to continue to 
investigate this negative customer behavior, predict credit score based on payment
frequency to reduce the incidence of non-payment of loans.



# Case Study 2: PEG Africa - Repayment

## Scenario 1: 
Given the DB write a SQL Query to get the amount paid by each customer, 
create a new column called "Total Amount Paid" and use this amount to create a column
called outstanding balance in which I will add the customer's outstanding balance after 
calculating "Total Amount Paid".

```{r}

# Question 1

# Loading the Dataset
db_1 <- read_excel(path = "Case 2 Data.xlsx", sheet = 1)


# Verifying Head Data
head(db_1)

# Converting Amount and Product_Price to Numeric
db_1$Amount <- as.numeric(db_1$Amount)
db_1$ProductPrice <- as.numeric(db_1$ProductPrice)


# Total Amount Paid by the Customer

total_amount_paid_by_Customer <- db_1 %>%
  group_by(CustomerId, ContractId, ProductPrice) %>%
  summarise(total_Payment = sum(Amount))

# Preview the result of the Query
head(total_amount_paid_by_Customer, 10)

# Clients with More than one Contract
customer_with_more_than_one_Contract <- total_amount_paid_by_Customer %>%
  # Filter by the customerID
  group_by(CustomerId) %>%
  # Count the number of times each Customer ID occurs
  summarise(freq = n()) %>%
  # Show Customers with more than one contract
  filter( freq > 1)

# Preview list of Customers with more than one Contract
head(customer_with_more_than_one_Contract, 10)


# Adding the Outstanding Balances

customer_Outstanding_Balance <- total_amount_paid_by_Customer %>%
  mutate(outstanding_Balance = ProductPrice  - total_Payment)

head(customer_Outstanding_Balance, 10)

```

Please Note that the Case Study was not clear on what happens to the Deposit of the Customer. The Query above totals all Amounts paid by the customer for each contract he/she holds without the Customer Down Payment.

If Customer Downpayment is added to amounts paid, the resulting Query is:

```{r}

# Now, Computing Total Amount Paid with Deposit inclusive

# Loading the Dataset
db_1 <- read_excel(path = "Case 2 Data.xlsx", sheet = 1)


# Verifying Head Data
head(db_1)

# Converting Amount and Product_Price to Numeric
db_1$Amount <- as.numeric(db_1$Amount)
db_1$ProductPrice <- as.numeric(db_1$ProductPrice)
db_1$Deposit <- as.numeric(db_1$Deposit)

total_amount_paid_by_Customer_with_Deposit <- db_1 %>%
  group_by(CustomerId, ContractId, ProductPrice, Deposit) %>%
  summarise(total_Payment = sum(Amount)) 


# Preview of Field
head(total_amount_paid_by_Customer_with_Deposit, 10)


# total_amount_paid_plus_Deposit = Deposit + Total_Paid_Amount
total_amount_paid_by_Customer_Plus_Deposit <- 
  total_amount_paid_by_Customer_with_Deposit %>%
  mutate(total_amount_paid_plus_Deposit = Deposit + total_Payment)


# Preview 
head(total_amount_paid_by_Customer_Plus_Deposit, 10)


# Outstanding Payments = Product_Price - (Total_Amount_Paid + Deposit) 

customer_Outstanding_Balance_after_Deposit_and_Tot_Payt <- total_amount_paid_by_Customer_Plus_Deposit %>%
  mutate(Outstanding_Balance_after_Deposit_and_Tot_Payt = ProductPrice - total_amount_paid_plus_Deposit)


# Preview
head(customer_Outstanding_Balance_after_Deposit_and_Tot_Payt[ , -3], 10)
```



## Scenario 2:

Write a query to get the sum_paid_to_date and arrears of each contractID at their
maximum date of activity.

```{r}
# Load the DB
db_2 <- read_excel(path = "Case 2 Data.xlsx", sheet = 2)


# Lets look at the Data

head(db_2)


# Getting the sum_paid_to_date and arrears at maximum date of activity

sum_Paid_to_Date <- db_2 %>%
  # Grouping by CustomerID and ContractID
  group_by(customerid, contractid) %>%
  # Filtering the Group by maximum activity date
  filter(date_of_activity == max(date_of_activity)) %>%  
  arrange(contractid)


head(sum_Paid_to_Date[ , -c(3)], 10)

```
  

# Case Study 3:

## Scenario: Demonstrating Business Understanding

For this section, I will break the corresponding KPI metrics into the following:

1. Overall: Total_Sales_by_Country, Product_Performance_per_country, Active_vs_Blocked_Customers_per_Country

2. Sales: Active_Contracts_per_SC, Blocked_Contracts_per_SC, Best_Performing_Products, Product_Sales_per_Country

3. Credit Management: Customer_performance_to_Date, Expected_DailyReceivable_vs_ActualReceived.

4. Finance: Expected_Accounts_Receivable_vs_Actual_Accounts_Receivable, ActualDaily_Deposits_vs_Target, Cash-to-Cash-Cycle.


## Group Level Data Analysis

Firstly, lets see which country is performing better:
```{r}
# Loading the dataset
db_3 <- read_excel("Case 3-Sample data.xlsx")

# Preview the data
head(db_3)

# Converting CustomerStatus to numeric
db_3$customerStatus <- ifelse(db_3$customerStatus == "Blocked", 0, 1)

# Converting Sum_Paid_to_Numeric
db_3$Sum_Paid_To_Date <- as.numeric(db_3$Sum_Paid_To_Date)


db_3 %>%
  group_by(CountryId) %>%
  summarise(Tot_Amt_Received_to_Date_loc_Curr = sum(Sum_Paid_To_Date)) %>%
  arrange(desc(Tot_Amt_Received_to_Date_loc_Curr))
```

Looks like PEG Africa's Senegal Office is doing something right! 


Next, we find out which Product is selling the most:

```{r}
db_3 %>%
  group_by(productTypeGeneral) %>%
  summarise(sum_paid_per_product = sum(Sum_Paid_To_Date)) %>%
  arrange(desc(sum_paid_per_product))
```

Now, how many Customers does each country have? Lets find out:

```{r}
db_3 %>%
  group_by(CountryId) %>%
  summarise(No_of_Customers = n())
```


Finally, lets find out the number of Active customers per group
```{r}
db_3 %>% 
  filter(customerStatus == 1) %>%
  group_by(CountryId) %>%
  summarise(Number_of_Active_Cust = sum(customerStatus))
```
It gets interesting as all customers in region CDI are blocked. What could be the
reason for this? Also, Senegal has more active clients than the Ghana office. Lets deep-dive into Senegal to find out what is hapenning in there.

## A Dive into Senegal

The best performing region is Senegal is: "SC Gandiaye":
```{r}
# Senegal Data
db_3 %>%
  filter(CountryId == "SN", customerStatus == 1) %>%
  group_by(`Sales location`) %>%
  summarise(Earnings_to_Date = sum(Sum_Paid_To_Date)) %>%
  arrange(desc(Earnings_to_Date))
```

Is there a repayment problem? Lets find this out in Senegal:
```{r}
# Repayment in Senegal
db_3 %>%
  filter(CountryId == "SN") %>%
  group_by(`Sales location`) %>%
  summarise(Repayment_rate_Perc = (mean(customerStatus) * 100)) %>%
  arrange(desc(Repayment_rate_Perc))
```

Surprisingly, there is no repayment problem in Senegal. I suggest that, management
at the group-level investigate strategies being implemented in Senegal's provinces
and repeat same at the other locations.

What product sells most in Senegal?
```{r}
# Best performing Products  in Senegal
db_3 %>%
  filter(CountryId == "SN") %>%
  group_by(productTypeGeneral) %>%
  summarise(Sales_by_Product = n()) %>%
  arrange(desc(Sales_by_Product))
```


## A look at Ghana

Lets take a look at the best performing Service Center in Ghana. Looks like Bogoso
leads the pack!
```{r}
db_3 %>%
  filter(CountryId == "GH", customerStatus == 1) %>%
  group_by(`Sales location`) %>%
  summarise(Earnings_to_Date = sum(Sum_Paid_To_Date)) %>%
  arrange(desc(Earnings_to_Date)) %>%
  head(10)
```

A look at the percentage of Active clients against total number of clients at the
service center (as a measure of repayment since defaulting customers are "blocked") reveals a staggering problem.

```{r}
# Repayment in Ghana
db_3 %>%
  filter(CountryId == "GH") %>%
  group_by(`Sales location`) %>%
  summarise(Repayment_rate_Perc = (mean(customerStatus) * 100)) %>%
  arrange(Repayment_rate_Perc) %>%
  head(10)
```

Ghanaian customers have a high default rate and are blocked due to non-repayment of loans. This is a revealing problem I have attempted to answer in case study 1.
Demographic data is needed on these customers to understand the attitude of the Ghanaian PEG-Africa customer.

What products sell most in Ghana?

```{r}
# Best performing Products  in Ghana
db_3 %>%
  filter(CountryId == "GH") %>%
  group_by(productTypeGeneral) %>%
  summarise(Sales_by_Product = n()) %>%
  arrange(desc(Sales_by_Product))
```


## Sample Visualization
```{r}
library(tidyverse)
library(bbplot)
require(scales)

db_3 %>%
  group_by(CountryId) %>%
  summarise(Tot_Amt_Received_to_Date_loc_Curr = sum(Sum_Paid_To_Date)) %>%
  arrange(desc(Tot_Amt_Received_to_Date_loc_Curr)) %>%
  ggplot(aes(reorder(x = CountryId, Tot_Amt_Received_to_Date_loc_Curr), y = Tot_Amt_Received_to_Date_loc_Curr, fill = CountryId)) +
  geom_col() +
  scale_y_continuous(labels = comma)+
  labs(title = "Earnings by Country",
       subtitle = "Country's Sales Performance", 
       x = "Earnings", y = "Country") + 
  bbc_style()
```

```{r}

db_3 %>%
  group_by(productTypeGeneral) %>%
  summarise(sum_paid_per_product = sum(Sum_Paid_To_Date)) %>%
  arrange(desc(sum_paid_per_product)) %>%
  ggplot(aes(reorder(x = productTypeGeneral, sum_paid_per_product), 
             y = sum_paid_per_product, fill = productTypeGeneral)) +
  geom_col() +
  scale_y_continuous(labels = comma) +
  scale_fill_brewer(palette = "Set1") +
  labs(title = "Performance of Product types", 
       subtitle = "Which product do customers prefer", x = "Product", 
       y = "Earnings by product") +
  bbc_style()
```


```{r}

db_3 %>%
  group_by(CountryId) %>%
  summarise(No_of_Customers = n()) %>%
  ggplot(aes(x = CountryId, y = No_of_Customers, fill = CountryId)) +
  geom_col() +
  scale_fill_brewer(palette = "Set1") +
  labs(title = "Number of Customers", subtitle = "How many Customers does your country have?", 
       x = "Country", y = "Number of Customer") +
  bbc_style()
```

```{r}
db_3 %>% 
  filter(customerStatus == 1) %>%
  group_by(CountryId) %>%
  summarise(Number_of_Active_Cust = sum(customerStatus)) %>%
  ggplot(aes(x = CountryId, y = Number_of_Active_Cust, fill = CountryId)) +
  geom_col() +
  scale_fill_brewer(palette = "Set1") +
  labs(title = "Number of Active Customers", subtitle = "How many Active Customers\n does your country have?", 
       x = "Country", y = "Number of Customer") +
  bbc_style()
```







